---
title: "2017_08_26 Lifehist glm"
author: "Virginia Chu"
date: "August 26, 2017"
output: html_document
---

```{r setup, include=FALSE}
#2017_08_26 glm attempt
setwd("C:/Users/virgc/Dropbox/Life history data")
lifedat<-read.csv("2017_09_19 Life history data.csv", header=TRUE)
#ruybal- dev is 1/time to adult

#

#lifedat$tempnum<-if (lifedat$Temp=="A") {20} else if (lifedat$Temp=="B") {24} else if (lifedat$Temp=="C") {28}

mod1<-glm(rate~Temp*Locality*Sex,data=lifedat,family=Gamma);summary(mod1)
mod2<-glm(rate~Temp,data=lifedat,family=Gamma);summary(mod2)
mod3<-glm(rate~Temp*Sex,data=lifedat,family=Gamma);summary(mod3)
mod4<-glm(rate~Temp*Locality,data=lifedat,family=Gamma);summary(mod4)
mod5<-glm(rate~Temp*Biome,data=lifedat,family=Gamma);summary(mod5)
mod6<-glm(rate~Temp*Biome*Sex,data=lifedat,family=Gamma);summary(mod6)
mod7<-glm(rate~Temp*State*Sex,data=lifedat,family=Gamma);summary(mod7)
mod8<-glm(rate~Temp*State,data=lifedat,family=Gamma);summary(mod8)
modelAIC<-AIC(mod1,mod2,mod3,mod4,mod5,mod6,mod7, mod8)
print(modelAIC)
print("mod1<-glm(rate~Temp*Locality*Sex,data=lifedat,family=Gamma);summary(mod1)")
print(summary(mod1))



#or this for automatic model building?
#makes a big model
lifebiglm <- lm(rate ~ (Family+Well)^2+(Family+Well)*(Temp+Locality+Sex), lifedat)
#steps through to remove parts to minimize AIC, trave=FALSE blocks model output clutter
lifesmallm <- step(lifebiglm,trace=FALSE)
#d
#drop1(lifesmallm,test="F")
#finalm <- lm(undercount~equip + econ  + perAA + equip:econ + equip:perAA, gavote)
#sumary(finalm)

#To try it out with binomial family instead of gamma
setwd("C:/Users/virgc/Dropbox/Life history data")
lifedat<-read.csv("2017_08_26 Life history data.csv", header=TRUE)
#ruybal- dev is 1/time to adult
lifedat$rate<- 1/lifedat$EmTime
#order by latitude
lifedat$Locality=factor(lifedat$Locality, c("ARS","APR","RPV","RMO","TLC","TPN","SJU"))
#
mod1b<-glm(rate~Temp*Locality*Sex,data=lifedat,family=binomial);summary(mod1b)
mod2b<-glm(rate~Temp,data=lifedat,family=binomial);summary(mod2b)
mod3b<-glm(rate~Temp*Sex,data=lifedat,family=binomial);summary(mod3b)
mod4b<-glm(rate~Temp*Locality,data=lifedat,family=binomial);summary(mod4b)
mod5b<-glm(rate~Temp*Biome,data=lifedat,family=binomial);summary(mod5b)
mod6b<-glm(rate~Temp*Biome*Sex,data=lifedat,family=binomial);summary(mod6b)
mod7b<-glm(rate~Temp*State*Sex,data=lifedat,family=binomial);summary(mod7b)
mod8b<-glm(rate~Temp*State,data=lifedat,family=binomial);summary(mod8b)
modelAICb<-AIC(mod1b,mod2b,mod3b,mod4b,mod5b,mod6b,mod7b, mod8b)
print(modelAICb)
print("mod1<-glm(rate~Temp*Locality*Sex,data=lifedat,family=Gamma);summary(mod1)")
print(summary(mod1))

#or this for automatic model building?
#makes a big model
lifebiglm <- lm(rate ~ (Family+Well)^2+(Family+Well)*(Temp+Locality+Sex), lifedat)
#steps through to remove parts to minimize AIC, trave=FALSE blocks model output clutter
lifesmallm <- step(lifebiglm,trace=FALSE)


#testing distribution
install.packages("car")
require (car)
install.packages("MASS")
require(MASS)

lifedat$Pupation.date<-lifedat$Pupation.date + 1
qqnorm(lifedat$Pupation.date.t, "norm")
qqp(lifedat$Pupation.date, "lnorm")


##testing with continuous temp rather than groups
mod1tn<-glm(rate~Tempnum*Locality*Sex,data=lifedat,family=Gamma);summary(mod1tn)
mod2tn<-glm(rate~Tempnum,data=lifedat,family=Gamma);summary(mod2tn)
mod3tn<-glm(rate~Tempnum*Sex,data=lifedat,family=Gamma);summary(mod3tn)
mod4tn<-glm(rate~Tempnum*Locality,data=lifedat,family=Gamma);summary(mod4tn)
mod5tn<-glm(rate~Tempnum*Biome,data=lifedat,family=Gamma);summary(mod5tn)
mod6tn<-glm(rate~Tempnum*Biome*Sex,data=lifedat,family=Gamma);summary(mod6tn)
mod7tn<-glm(rate~Tempnum*State*Sex,data=lifedat,family=Gamma);summary(mod7tn)
mod8tn<-glm(rate~Tempnum*State,data=lifedat,family=Gamma);summary(mod8tn)
modelAICtn<-AIC(mod1tn,mod2tn,mod3tn,mod4tn,mod5tn,mod6tn,mod7tn, mod8tn)
print(modelAICtn)
print("mod4tn<-glm(rate~Tempnum*Locality,data=lifedat,family=Gamma);summary(mod4tn)")
print(summary(mod4tn))

```

## R Markdown

# Description of columns:
## Independent variables
* Biome [Amazon, Cerrado, Mata Atlantica]- Categorical/nominal
    + State [Amazonas, Rondonia, Tocantins, Mata Atlantica]-Ordinal
    + Locality [ARS, APR, RPV, RMO, TLC, TPN, SJU]-Ordinal
    + Latitude [-2.864, -3.028, -8.742, -9.223, -10.7, -10.796, -22.611 ]- Ordinal
* Family [for lab reared, which female within the biome/locality they came from]- Categorical
  + Well grouping [1,2,3]- Ordinal (5 larvae raised in each of 3 wells per female per temperature)
* Temperature [ 20 (A), 24 (B), 28C (C)]- Ordinal
* Sex [M/F]- Categorical

## Dependent variables
* Death status (0,1,2) - Ordinal 0=died as larvae, 1= died as pupae, 2=died as adult
* LL- Larvae life length (days)- Discrete, all
* sLL- Larvae life length (days)- Discrete, only ones that went to adult
* PL- Pupal life length (days)- Discrete (Emergence.date-pupation.date)
* AL- Adult life length (days)- Discrete (Deathtime-Emergence.date)
* EmTime- time to emergence (Emergence.date-Hatch.day)
* Wing length (mm)- Continuous
* Rate of development
* Sex ratio



```{r cars}
print("Gamma distribution results")
print(modelAIC)
print("mod1<-glm(rate~Temp*Locality*Sex,data=lifedat,family=Gamma);summary(mod1)")
print(summary(mod1))

print("Binomial distribution results")
print(modelAICb)
print("mod2b<-glm(rate~Temp,data=lifedat,family=binomial);summary(mod2b)")
print(summary(mod2b))

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
