---
title: "2017_10_17 Life hist glm data analysis"
author: "Virginia Chu"
date: "October 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
##work comp
setwd("C:/Users/vmc04/Documents/GitHub/wingproj/lifehist")
lifehist<-read.csv("2017_10_17 Life history data new clean file.csv", header=TRUE)
lifehist$Locality=factor(lifehist$Locality, c("ARS","APR","RPV","RMO","TLC","TPN","SJU"))

#print(head(lifehist))
##libraries needed
library(ggplot2)
library(lme4)

## glm for rate of development
#ruybal rate= 1/time to adult [AdRate]
mod1<-glm(AdRate~Temp_num*Locality*Sex,data=lifehist,family=Gamma);summary(mod1)
mod2<-glm(AdRate~Temp_num,data=lifehist,family=Gamma);summary(mod2)
mod3<-glm(AdRate~Temp_num*Sex,data=lifehist,family=Gamma);summary(mod3)
mod4<-glm(AdRate~Temp_num*Locality,data=lifehist,family=Gamma);summary(mod4)
mod5<-glm(AdRate~Temp_num*Biome,data=lifehist,family=Gamma);summary(mod5)
mod6<-glm(AdRate~Temp_num*Biome*Sex,data=lifehist,family=Gamma);summary(mod6)
mod7<-glm(AdRate~Temp_num*State*Sex,data=lifehist,family=Gamma);summary(mod7)
mod8<-glm(AdRate~Temp_num*State,data=lifehist,family=Gamma);summary(mod8)
modelAIC<-AIC(mod1,mod2,mod3,mod4,mod5,mod6,mod7, mod8)
print(modelAIC)
write.table(modelAIC, "clipboard", sep="/t")
print("mod1<-glm(rate~Temp*Locality*Sex,data=lifedat,family=Gamma);summary(mod1)")
#print(summary(mod1))


mod1.1<-glm(AdRate~Temp_num^2*Locality*Sex,data=lifehist,family=Gamma);summary(mod1.1)
mod2.1<-glm(AdRate~Temp_num^2,data=lifehist,family=Gamma);summary(mod2.1)
mod3.1<-glm(AdRate~Temp_num^2*Sex,data=lifehist,family=Gamma);summary(mod3.1)
mod4.1<-glm(AdRate~Temp_num^2*Locality,data=lifehist,family=Gamma);summary(mod4.1)
mod5.1<-glm(AdRate~Temp_num^2*Biome,data=lifehist,family=Gamma);summary(mod5.1)
mod6.1<-glm(AdRate~Temp_num^2*Biome*Sex,data=lifehist,family=Gamma);summary(mod6.1)
mod7.1<-glm(AdRate~Temp_num^2*State*Sex,data=lifehist,family=Gamma);summary(mod7.1)
mod8.1<-glm(AdRate~Temp_num^2*State,data=lifehist,family=Gamma);summary(mod8.1)
modelAIC1.1<-AIC(mod1.1,mod2.1,mod3.1,mod4.1,mod5.1,mod6.1,mod7.1, mod8.1)
print(modelAIC1.1)
#print("mod1<-glm(rate~Temp*Locality*Sex,data=lifedat,family=Gamma);summary(mod1)")

mod1.2<-glm(AdRate~Temp_let*Locality*Sex,data=lifehist,family=Gamma);summary(mod1.2)
mod2.2<-glm(AdRate~Temp_let,data=lifehist,family=Gamma);summary(mod2.2)
mod3.2<-glm(AdRate~Temp_let*Sex,data=lifehist,family=Gamma);summary(mod3.2)
mod4.2<-glm(AdRate~Temp_let*Locality,data=lifehist,family=Gamma);summary(mod4.2)
mod5.2<-glm(AdRate~Temp_let*Biome,data=lifehist,family=Gamma);summary(mod5.2)
mod6.2<-glm(AdRate~Temp_let*Biome*Sex,data=lifehist,family=Gamma);summary(mod6.2)
mod7.2<-glm(AdRate~Temp_let*State*Sex,data=lifehist,family=Gamma);summary(mod7.2)
mod8.2<-glm(AdRate~Temp_let*State,data=lifehist,family=Gamma);summary(mod8.2)
modelAIC1.2<-AIC(mod1.2,mod2.2,mod3.2,mod4.2,mod5.2,mod6.2,mod7.2, mod8.2)
print(modelAIC1.2)

mod1.3<-glm(AdRate~Temp_num*Locality*Sex,data=lifehist,family=binomial);summary(mod1.3)
mod2.3<-glm(AdRate~Temp_num,data=lifehist,family=binomial);summary(mod2.3)
mod3.3<-glm(AdRate~Temp_num*Sex,data=lifehist,family=binomial);summary(mod3.3)
mod4.3<-glm(AdRate~Temp_num*Locality,data=lifehist,family=binomial);summary(mod4.3)
mod5.3<-glm(AdRate~Temp_num*Biome,data=lifehist,family=binomial);summary(mod5.3)
mod6.3<-glm(AdRate~Temp_num*Biome*Sex,data=lifehist,family=binomial);summary(mod6.3)
mod7.3<-glm(AdRate~Temp_num*State*Sex,data=lifehist,family=binomial);summary(mod7.3)
mod8.3<-glm(AdRate~Temp_num*State,data=lifehist,family=binomial);summary(mod8.3)
modelAIC1.3<-AIC(mod1.3,mod2.3,mod3.3,mod4.3,mod5.3,mod6.3,mod7.3, mod8.3)
print(modelAIC1.3)

mod1.4<-glm(AdRate~Temp_num*Locality*Sex,data=lifehist,family=binomial(logit));summary(mod1.4)
mod2.4<-glm(AdRate~Temp_num,data=lifehist,family=binomial(logit));summary(mod2.4)
mod3.4<-glm(AdRate~Temp_num*Sex,data=lifehist,family=binomial(logit));summary(mod3.4)
mod4.4<-glm(AdRate~Temp_num*Locality,data=lifehist,family=binomial(logit));summary(mod4.4)
mod5.4<-glm(AdRate~Temp_num*Biome,data=lifehist,family=binomial(logit));summary(mod5.4)
mod6.4<-glm(AdRate~Temp_num*Biome*Sex,data=lifehist,family=binomial(logit));summary(mod6.4)
mod7.4<-glm(AdRate~Temp_num*State*Sex,data=lifehist,family=binomial(logit));summary(mod7.4)
mod8.4<-glm(AdRate~Temp_num*State,data=lifehist,family=binomial(logit));summary(mod8.4)
modelAIC1.4<-AIC(mod1.4,mod2.4,mod3.4,mod4.4,mod5.4,mod6.4,mod7.4, mod8.4)
print(modelAIC1.4)

##subset data by sexes
lifehist_females<-subset(lifehist, Sex=="F", select=c(1:33))
lifehist_males<-subset(lifehist, Sex=="M", select=c(1:33))

mod1.5<-glm(AdRate~Temp_let*Locality,data=lifehist_females,family=binomial(logit));summary(mod1.5)
mod2.5<-glm(AdRate~Temp_let,data=lifehist_females,family=binomial(logit));summary(mod2.5)
mod3.5<-glm(AdRate~Locality,data=lifehist_females,family=binomial(logit));summary(mod3.5)
mod4.5<-glm(AdRate~Temp_let*Biome,data=lifehist_females,family=binomial(logit));summary(mod4.5)
mod5.5<-glm(AdRate~Temp_let*State,data=lifehist_females,family=binomial(logit));summary(mod5.5)
mod6.5<-glm(AdRate~Biome,data=lifehist_females,family=binomial(logit));summary(mod6.5)
mod7.5<-glm(AdRate~State,data=lifehist_females,family=binomial(logit));summary(mod7.5)
modelAIC1.5<-AIC(mod1.5,mod2.5,mod3.5,mod4.5,mod5.5,mod6.5,mod7.5)
print(modelAIC1.5)

mod1.6<-glm(AdRate~Temp_num*Locality,data=lifehist_females,family=binomial(logit));summary(mod1.6)
mod2.6<-glm(AdRate~Temp_num,data=lifehist_females,family=binomial(logit));summary(mod2.6)
mod3.6<-glm(AdRate~Locality,data=lifehist_females,family=binomial(logit));summary(mod3.6)
mod4.6<-glm(AdRate~Temp_num*Biome,data=lifehist_females,family=binomial(logit));summary(mod4.6)
mod5.6<-glm(AdRate~Temp_num*State,data=lifehist_females,family=binomial(logit));summary(mod5.6)
mod6.6<-glm(AdRate~Biome,data=lifehist_females,family=binomial(logit));summary(mod6.6)
mod7.6<-glm(AdRate~State,data=lifehist_females,family=binomial(logit));summary(mod7.6)
modelAIC1.6<-AIC(mod1.6,mod2.6,mod3.6,mod4.6,mod5.6,mod6.6,mod7.6)
print(modelAIC1.6)
```

## R Markdown

Life history data
# Description of columns:
## Independent variables
* Biome [Amazon, Cerrado, Mata Atlantica]- Categorical/nominal
    + State [Amazonas, Rondonia, Tocantins, Mata Atlantica]-Ordinal
    + Locality [ARS, APR, RPV, RMO, TLC, TPN, SJU]-Ordinal
    + Latitude [-2.864, -3.028, -8.742, -9.223, -10.7, -10.796, -22.611 ]- Ordinal
* Family [for lab reared, which female within the biome/locality they came from]- Categorical
  + Well grouping [1,2,3]- Ordinal (5 larvae raised in each of 3 wells per female per temperature)
* Temperature [ 20 (A), 24 (B), 28C (C)]- Ordinal
* Temp_num
* Sex [M/F]- Categorical

## Dependent variables
* Death status (0,1,2) - Ordinal 0=died as larvae, 1= died as pupae, 2=died as adult
* LL- Larvae life length (days)- Discrete, all
* sLL- Larvae life length (days)- Discrete, only ones that went to adult
* PL- Pupal life length (days)- Discrete (Emergence.date-pupation.date)
* AL- Adult life length (days)- Discrete (Deathtime-Emergence.date)
* EmTime- time to emergence (Emergence.date-Hatch.day)
* Wing length (mm)- Continuous
* Rate of development
* Sex ratio

##Datafile column headers
ID	Biome	State	Latitude	Locality	Family	Temp_let	Temp_num	Well	Hatch.day	LD	Pupation.date	PD	Emergence.date	Adult.death	Sex	Wing.length..mm.	Grouping	Plate.info	sLL	Emtime	Preadult.death	Sex1	Death_stat	AdRate	LarvRate	Puplen	Puprate	D.pread.rate	AL


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
