---
title: "Life hist Cox regression"
author: "Virginia Chu"
date: "August 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
#home laptop
setwd("C:/Users/virgc/Dropbox/Life history data")

#work comp
setwd("C:/Users/vmc04/Dropbox/Life history data")
#file
lifedat<-read.csv("2017_08_26 Life history data.csv", header=TRUE)
#order by latitude
lifedat$Locality=factor(lifedat$Locality, c("ARS","APR","RPV","RMO","TLC","TPN","SJU"))

#survival package
install.packages("survival")
library(survival)

x1<-aggregate(lifedata$AL, list(temp=lifedat$Temp, loc=lifedat$Locality,sex=lifedat$Sex),length)
write.table(x1, "clipboard", sep='\t', row.names=FALSE)
```

## R Markdown
Sources used- http://rstudio-pubs-static.s3.amazonaws.com/5896_8f0fed2ccbbd42489276e554a05af87e.html


Description of columns:
.	Biome [Amazon, Cerrado, Mata Atlantica]- Categorical/nominal
 o	State [Amazonas, Rondonia, Tocantins, Mata Atlantica]-Ordinal
 o	Locality [ARS, APR, RPV, RMO, TLC, TPN, SJU]-Ordinal
 o	Latitude [-2.864, -3.028, -8.742, -9.223, -10.7, -10.796, -22.611 ]- Ordinal
.	Family [for lab reared, which female within the biome/locality they came from]- Categorical
 o	Well grouping [1,2,3]- Ordinal (5 larvae raised in each of 3 wells per female per temperature)
.	Temperature [ 20 (A), 24 (B), 28C (C)]- Ordinal
.	Sex [M/F]- Categorical

.	Death status (0,1,2) - Ordinal 0=died as larvae, 1= died as pupae, 2=died as adult
.	LL- Larvae life length (days)- Discrete, all
.	sLL- Larvae life length (days)- Discrete, only ones that went to adult
.	PL- Pupal life length (days)- Discrete (Emergence.date-pupation.date)
.	AL- Adult life length (days)- Discrete (Deathtime-Emergence.date)
.	EmTime- time to emergence (Emergence.date-Hatch.day)
.	Wing length (mm)- Continuous
.	Rate of development

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
