---
title: "Life history An darlingi Brazil Statistical analysis 11/2/2017"
author: "Virginia Chu"
date: "November 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(plyr)
library (survival)
library (survminer)
library(lme4)
library(ggplot2)
setwd("C:/Users/vmc04/Documents/GitHub/wingproj/lifehist")
#laptop
#setwd("C:/Users/virgc/GitHub/wingproj/lifehist")
#Read in datafile
lifehistshort<-read.csv("2017_10_18 Life history shortened.csv", header=TRUE)
#Factor levels of locality, temp num, and lat_group
lifehistshort$Locality=factor(lifehistshort$Locality, c("ARS","APR","RPV","RMO","TLC","TPN","SJU"))
lifehistshort$Lat_group=factor(lifehistshort$Lat_group, c(1,2,3))
lifehistshort$Temp_num=factor(lifehistshort$Lat_group, c(20,24,28))

## Effect of temp on proportion surviving by sex and location
lifetot<-read.csv("2017_11_21 Life hist prop temp.csv", header=TRUE)
lifecont1<-read.csv("2017_10_19 Life hist totals for proportion  survived from temp1.csv",header=TRUE)

##Descriptive tables

#LarvRate
larvae<- ddply(lifehistshort,c("Lat_group", "Temp_let"), summarise,
               N= length(LarvRate),
               mean=mean(LarvRate, na.rm=TRUE),
               sd= sd(LarvRate, na.rm=TRUE),
               se=sd/sqrt(N))
Preadtime<-ddply(lifehistshort,c("Lat_group", "Temp_let"), summarise,
               N= length(AdRate),
               mean=mean(AdRate, na.rm=TRUE),
               sd= sd(AdRate, na.rm=TRUE),
               se=sd/sqrt(N))
Emergtime<-ddply(lifehistshort,c("Lat_group", "Temp_let"), summarise,
               N= length(Emtime),
               mean=mean(Emtime, na.rm=TRUE),
               sd= sd(Emtime, na.rm=TRUE),
               se=sd/sqrt(N))
Adlife<-ddply(lifehistshort,c("Lat_group", "Temp_let"), summarise,
               N= length(AL),
               mean=mean(AL, na.rm=TRUE),
               sd= sd(AL, na.rm=TRUE),
               se=sd/sqrt(N))
Winglength<-ddply(lifehistshort,c("Lat_group", "Temp_let"), summarise,
               N= length(Wing.length..mm.),
               mean=mean(Wing.length..mm., na.rm=TRUE),
               sd= sd(Wing.length..mm., na.rm=TRUE),
               se=sd/sqrt(N))

#1a. temp had significant affect? p=<2.2e-16

table( lifehistshort$Temp_let, lifehistshort$Death_stat)
prop.test(table(lifehistshort$Temp_let, lifehistshort$Death_stat), correct=FALSE)


#1b. locality had significant affect? p=<2.2e-16
table( lifehistshort$Locality, lifehistshort$Death_stat)
prop.test(table(lifehistshort$Locality, lifehistshort$Death_stat), correct=FALSE)

#1c. Association between temp and location interaction  and proportion survived to adulthood, p=0.4249
#Used mantelhaen.test
data=mutate(lifecont1, Locality=factor(Locality, levels=unique(Locality)),
            Temperature=factor(Temperature, levels=unique(Temperature)),
            Status=factor(Status, levels=unique(Status)))

data.xtabs=xtabs(Count~Locality+Temperature+Status, data=data)
ftable(data.xtabs)
mantelhaen.test(data.xtabs)

#1d. Cox regression?
cox_propt<-coxph(Surv(time, Death_stat)~ Temp_let+ Locality+Temp_let:Locality , lifehistshort)
summary(cox_propt)
cox_propl<-coxph(Surv(time, Death_stat)~ Temp_let +Lat_group+Temp_let:Lat_group, lifehistshort)
summary(cox_propl)


#1e. Log rank test
sd_t<- survdiff(Surv(time, Death_stat)~ Temp_let, lifehistshort)
sd_l<- survdiff(Surv(time, Death_stat)~ Locality, lifehistshort)
sd_g<- survdiff(Surv(time, Death_stat)~ Lat_group, lifehistshort)

#2a. Association between temp and proportion survived by sex, p=0.347
table( lifehistshort$Temp_let, lifehistshort$Sex1)
prop.test(table(lifehistshort$Temp_let, lifehistshort$Sex1), correct=FALSE)
#2b. Association between location and proportion survived by sex, p=0.4634
table( lifehistshort$Locality, lifehistshort$Sex1)
prop.test(table(lifehistshort$Locality, lifehistshort$Sex1), correct=FALSE)

#2c. Association between temp and location interaction  and proportion survived to adulthood, p=0.4335
#Used mantelhaen.test
data1=mutate(lifetot, Locality=factor(Locality, levels=unique(Locality)),
            Temperature=factor(Temp, levels=unique(Temp)),
            Status=factor(Status, levels=unique(Status)))

data1.xtabs=xtabs(Count~Locality+Temperature+Status, data=data1)
ftable(data1.xtabs)
mantelhaen.test(data1.xtabs)

#3a. GLMM with gamma
lr_fit1g<-glmer(LarvRate~(1|Biome)+(1|State) +(1|Family), data=lifehistshort, family=Gamma)
summary(lr_fit1g)
lr_fit2g<-glmer(LarvRate~Lat_group+(1|Biome)+(1|State), data=lifehistshort, family=Gamma)
summary(lr_fit2g)
lr_fit3g<-glmer(LarvRate~Lat_group+Temp_let+(1|Biome)+(1|State), data=lifehistshort, family=Gamma)
summary(lr_fit3g)
lr_fit4g<-glmer(LarvRate~(1|Biome)+(1|State)+Temp_let*Lat_group, data=lifehistshort, family=Gamma)
summary(lr_fit4g)

lrfitb_modg<-AIC(lr_fit1g, lr_fit2g, lr_fit3g, lr_fit4g)
lrfitb_modg

#3b. GLMM with gamma by sex
lr_fit6g<-glmer(LarvRate~(1|Biome)+(1|State)+Lat_group+Temp_let+Sex, data=lifehistshort, family=Gamma)
summary(lr_fit6g)
lr_fit7g<-glmer(LarvRate~(1|Biome)+(1|State)+Lat_group*Temp_let*Sex, data=lifehistshort, family=Gamma)
summary(lr_fit7g)

lrfitb_modg1<-AIC(lr_fit6g, lr_fit7g)
lrfitb_modg1

#3c. GLMM with gamma by sex for preadult development
pr_fit1<- glmer(AdRate~ (1|Biome)+(1|State)+Temp_let*Lat_group*Sex, data=lifehistshort, family=Gamma)
summary(pr_fit1)


#4a. GLMM with gamma for preadult development
pr_fit1<- glmer(AdRate~ (1|Biome)+(1|State)+Temp_let*Lat_group*Sex, data=lifehistshort, family=Gamma)
summary(pr_fit1)

```

## Life history data description
# Description of columns:
## Independent variables
* Biome [Amazon, Cerrado, Mata Atlantica]- Categorical/nominal
    + State [Amazonas, Rondonia, Tocantins, Mata Atlantica]-Ordinal
    + Locality [ARS, APR, RPV, RMO, TLC, TPN, SJU]-Ordinal
    + Latitude [-2.864, -3.028, -8.742, -9.223, -10.7, -10.796, -22.611 ]- Ordinal
* Lat_group- 1 (low, near equator), 2 (mid), 3 (high, near Rio))
* Family [for lab reared, which female within the biome/locality they came from]- Categorical
 (5 larvae raised in each of 3 wells per female per temperature)
* Temp_let- A, B, C
* Temp_num- 20,24,28
* Sex [M/F]- Only from surviving adults, did not sex pre-adults
* Sex1 [1/0]- 0-Female, 1-Male, 

## Dependent variables
* Wing length (mm)
* sLL- Larvae life length (days)- Discrete, only ones that went to adult
* dLL- Larvae life length (days)- Discrete, length of life of preadults that didn't develop to adults
* dPL- Pupal+larvae life length (days)- Length of time that pupae that died, survived 
* EmTime- time to emergence (Emergence.date-Hatch.day)
* Preadult.death- Time of life (in days) that preadults (that didn't make it to adulthood)
* Death_stat [0/1]- 0- preadult, 1-adult death status
* AdRate- Rate of development of mosquitoes that survived to adulthood
* LarvRate- Rate of development of 
* Puplen- Length in days of pupal life
* Puprate- Rate of development of pupae
* D.pread.rate- rate of development of mosquitoes that didn't survive to adulthood
* AL- Adult life length (days)- Discrete (Deathtime-Emergence.date)
* time- total length of time alive, regardless of death status as adult or larvae


##Datafile column headers
 ID  Biome    State Latitude Lat_group Locality Family Temp_let Temp_num Sex Sex1 Wing.length..mm. sLL dLL dPL Emtime
  Preadult.death Death_stat     AdRate   LarvRate Puplen   Puprate D.pread.rate AL time
  
##Plots


## Questions and Analysis

1. Does temperature or population location affect the proportion that survived to adulthood?

* Surviving proportion based on death as adult vs. preadult
Proportions test: prop.test and mantel.haen

1a. Temperature alone : p<2.2e-16

1b. Locality alone : p<2.2e-16

1c. Temperature and Locality :  p=0.4249

>The combined effects of temperature and locality did not have a significant effect on the proportion that survived to adulthood.

2. Does temperature or population location affect the proportion that survived to adulthood by sex?
* Surviving sex ratio
Proportions test: prop.test and mantel.haen

2a. Temperature alone : p=0.347

2b. Locality alone : p=0.4634

2c. Temperature and Locality :  p=0.4335

>The combined effects of temperature and locality did not have a significant effect on the proportion by sex that survived to adulthood.

3. Does temperature or population location affect the rate of larvae alone or preadult (larvae and pupae) development? 

3a. GLMM, gamma dist
Larvae rate ~ (1 | Biome) + (1 | State) + Temp_let * Lat_group
"lr_fit4g
(Intercept)           19.5005     0.5503   35.44  < 2e-16 ***
Temp_letB             -2.6475     0.1665  -15.90  < 2e-16 ***
Temp_letC             -4.6888     0.1598  -29.34  < 2e-16 ***
Lat_group2             2.1412     0.1761   12.16  < 2e-16 ***
Lat_group3             3.7665     0.8099    4.65 3.31e-06 ***
Temp_letB:Lat_group2  -0.7457     0.2263   -3.29 0.000985 ***
Temp_letC:Lat_group2  -0.3757     0.2220   -1.69 0.090590 .  
Temp_letB:Lat_group3  -1.7524     0.3765   -4.65 3.25e-06 ***
Temp_letC:Lat_group3  -2.4423     0.4688   -5.21 1.90e-07 ***"

>Stronger differences between all groups at 20C and 24C, 28C has mid-latitude similar to low. Low lat (Lat_group =1) fastest development at all temperatures

3b.GLMM with gamma by sex
LarvRate~(1|Biome)+(1|State)+Lat_group*Temp_let*Sex
"            Estimate Std. Error t value Pr(>|z|)    
(Intercept) 18.54714    1.90907    9.72  < 2e-16 ***
Lat_group2   1.71167    0.09297   18.41  < 2e-16 ***
Lat_group3   2.48619    0.79110    3.14  0.00167 ** 
Temp_letB   -3.18989    0.10748  -29.68  < 2e-16 ***
Temp_letC   -5.03073    0.10801  -46.58  < 2e-16 ***
SexF         1.49009    1.81701    0.82  0.41217    
SexM         1.12456    1.81678    0.62  0.53593"  

>Sex did not affect rate of larvae development.

3c. GLMM of preadult rate by sex
AdRate~ (1|Biome)+(1|State)+Temp_let*Lat_group*Sex

>Sex did not affect rate of preadult development.

4. Does temperature or population location affect the adult life span? 


```{}

```

## Plots

```{r pressure, echo=FALSE}
ggplot(larvae, aes(x=Temp_let, y=mean, fill=Lat_group)) + geom_bar(position=position_dodge(), stat="identity")+ geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd),width=.4,position=position_dodge(.9)) +xlab("Latitude group")+ ylab("Rate of Larvae Development (1/day)")+ ggtitle("Larvae rate of development by temperature and latitude group (SD bars)")

ggplot(Preadtime, aes(x=Temp_let, y=mean, fill=Lat_group)) + geom_bar(position=position_dodge(), stat="identity")+ geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd),width=.4,position=position_dodge(.9)) +xlab("Latitude group")+ ylab("Rate of Preadult Development (1/day)")+ ggtitle("Preadult rate of development by temperature and latitude group (SD bars)")

ggplot(Emergtime, aes(x=Temp_let, y=mean, fill=Lat_group)) + geom_bar(position=position_dodge(), stat="identity")+ geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd),width=.4,position=position_dodge(.9)) +xlab("Latitude group")+ ylab("Time to eclosion (days)")+ ggtitle("Time to eclosion by temperature and latitude group (SD bars)")

ggplot(Adlife, aes(x=Temp_let, y=mean, fill=Lat_group)) + geom_bar(position=position_dodge(), stat="identity")+ geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd),width=.4,position=position_dodge(.9)) +xlab("Latitude group")+ ylab("Length of Adult life (days)")+ ggtitle("Length of adult life by temperature and latitude group (SD bars)")

ggplot(Winglength, aes(x=Temp_let, y=mean, fill=Lat_group)) + geom_bar(position=position_dodge(), stat="identity")+ geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd),width=.4,position=position_dodge(.9)) +xlab("Latitude group")+ ylab("Wing length (mm)")+ ggtitle("Wing length by temperature and latitude group (SD bars)")
```
