---
title: "Life history An darlingi Brazil Statistical analysis 11/2/2017"
author: "Virginia Chu"
date: "November 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C:/Users/vmc04/Documents/GitHub/wingproj/lifehist")

lifehistshort<-read.csv("2017_10_18 Life history shortened.csv", header=TRUE)
lifehistshort$Locality=factor(lifehistshort$Locality, c("ARS","APR","RPV","RMO","TLC","TPN","SJU"))
lifehistshort$Lat_group=factor(lifehistshort$Lat_group, c(1,2,3))
```

## Life history data description
# Description of columns:
## Independent variables
* Biome [Amazon, Cerrado, Mata Atlantica]- Categorical/nominal
    + State [Amazonas, Rondonia, Tocantins, Mata Atlantica]-Ordinal
    + Locality [ARS, APR, RPV, RMO, TLC, TPN, SJU]-Ordinal
    + Latitude [-2.864, -3.028, -8.742, -9.223, -10.7, -10.796, -22.611 ]- Ordinal
* Family [for lab reared, which female within the biome/locality they came from]- Categorical
 (5 larvae raised in each of 3 wells per female per temperature)
* Temp_let- A, B, C
* Temp_num- 20,24,28
* Sex [M/F]- Only from surviving adults, did not sex pre-adults
* Sex1 [1/0]- 0-Female, 1-Male, 

## Dependent variables
* Wing length (mm)
* sLL- Larvae life length (days)- Discrete, only ones that went to adult
* dLL- Larvae life length (days)- Discrete, length of life of preadults that didn't develop to adults
* dPL- Pupal+larvae life length (days)- Length of time that pupae that died, survived 
* EmTime- time to emergence (Emergence.date-Hatch.day)
* Preadult.death- Time of life (in days) that preadults (that didn't make it to adulthood)
* Death_stat [0/1]- 0- preadult, 1-adult death status
* AdRate- 
* LarvRate- 
* Puplen-
* Puprate-
* D.pread.rate-
* AL- Adult life length (days)- Discrete (Deathtime-Emergence.date)
* time-


* Rate of development
* Sex ratio

##Datafile column headers
 ID  Biome    State Latitude Lat_group Locality Family Temp_let Temp_num Sex Sex1 Wing.length..mm. sLL dLL dPL Emtime
  Preadult.death Death_stat     AdRate   LarvRate Puplen   Puprate D.pread.rate AL time

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
